Implementation and Evaluation of the Simple Network Management Protocol over IEEE 802.15.4 Radios under the Contiki Operating System\\ MSc. Thesis Proposal

### 2010-01-31 trunk/doc/msc-proposal/msc-snmp.tex
proposal: SNMP security models

@@

There are several types of defined applications (Figure \ref{fig:snmparch}), such as a command generator which monitor and manipulate management data, a command responder providing access to management data, a notification originator initiating asynchronous messages, a notification receiver processing these messages, and a proxy forwarder which forwards messages between entities. Applications may use the services provided by the SNMP engine. An SNMP entity which includes one or more command generator and notification receiver applications is called an SNMP manager. An entity containing one or more command generator and notification receiver applications is called an SNMP manager.

\section{Message Format}\label{sec:messformat}
The SNMPv3 message (Figure \ref{fig:snmp-format}) contains the SNMP version, fields for global data (such as the message identifier, the maximum message size, the security model and the level of security), fields for the security model information and naming scope (context identifier and name), and finally the protocol data unit (PDU).
\begin{figure}[htp]	
\begin{center}
    \includegraphics[scale = 0.60]{img/snmpv3-message.pdf}
    \caption{SNMPv3 message format}   
	\label{fig:snmp-format}
\end{center}
\end{figure}

@@

\section{Protocol Operations}
As described in Section \ref{sec:messformat}, an SNMP message encapsulates a PDU which specifies the operations performed by the receiving SNMP engine. The SNMP protocol defines PDUs for the following operations: Get, GetNext, GetBulk, Response, Set, Trap, Inform.

In addition, the Report protocol operation can be used internally to allow SNMP engines to communicate with each other for error notifications and clock synchronization.

\section{Security Models}

>>>>

\section{Management Information Base}
SNMP itself does not specify which information is offered by a managed system. Management information which can be accessed via SNMP is viewed as a collection of managed objects in a virtual information store, called the Management Information Base (MIB). MIBs describe the structure of the device's management information using a hierarchical namespace containing object identifiers (OID). Each OID identifies a variable that can be read or set via SNMP. Collections of related objects are defined in MIB modules written in the Structure of Management Information (SMI) data definition language, which is an adapted subset of the Abstract Syntax Notation One (ASN.1) language. A number of specific MIBs has been defined as part of the SNMP suite specifying managed objects for the most common network management subjects.

@@

There are several types of defined applications (Figure \ref{fig:snmparch}), such as a command generator which monitor and manipulate management data, a command responder providing access to management data, a notification originator initiating asynchronous messages, a notification receiver processing these messages, and a proxy forwarder which forwards messages between entities. Applications may use the services provided by the SNMP engine. An SNMP entity which includes one or more command generator and notification receiver applications is called an SNMP manager. An SNMP entity containing one or more command responder and notification originator applications is called an SNMP agent.

\section{Message Format}\label{sec:messformat}
The SNMPv3 message (Figure \ref{fig:snmp-format}) contains the SNMP version, fields for global data (such as the message identifier, the maximum message size, the security model and the level of security), fields for the security model information and naming scope (context identifier and name), and finally the protocol data unit (PDU).
\begin{figure}[htp]	
\begin{center}
    \includegraphics[scale = 0.60]{img/snmpv3-message.pdf}
    \caption{SNMPv3 Message Format}   
	\label{fig:snmp-format}
\end{center}
\end{figure}

@@

The common format of protocol data units is shown in Figure \ref{fig:snmp-pdu}. This format is used for all operations described in Section \ref{sec:opetations}, except for the GetBulk request. The PDU type field has an integer value which identifies the type of the PDU. The request identifier is used to match requests with replies. It is generated by the device that sends a request and copied into this field in a response PDU by the responding SNMP entity. The error status tells the requesting SNMP entity the result of its request. A value of zero indicates that no error occurred; the other values indicate what sort of error happened. When error status is non-zero, the error index field points to the object generated the error, and has the value of zero in a request. The variable bindings is a set of name-value pairs identifying the MIB objects in the PDU, and in the case of messages other than requests, containing their values.
\begin{figure}[htp]	
\begin{center}
    \includegraphics[scale = 0.60]{img/snmp-pdu.pdf}
    \caption{SNMP Common PDU Format}   
	\label{fig:snmp-pdu}
\end{center}
\end{figure}

\section{Protocol Operations}\label{sec:opetations}
As described in Section \ref{sec:messformat}, an SNMP message encapsulates a PDU which specifies the operations performed by the receiving SNMP engine. The SNMP protocol defines PDU formats for the following operations: \texttt{Get}, \texttt{GetNext}, \texttt{GetBulk}, \texttt{Response}, \texttt{Set}, \texttt{Trap}, \texttt{Inform} and \texttt{Report} (Figure \ref{fig:snmp-operations}).

\begin{figure}[htp]	
\begin{center}
	\label{fig:snmp-operations}
    \includegraphics[scale = 0.80]{img/snmp-operations.pdf}
    \caption{SNMP Operations}   
\end{center}
\end{figure}

The \texttt{Get} request initiated by a manager allows to access and retrieve the value of one or more instances of management information. The agent processes each variable binding in the variable-binding list of the received \texttt{Get} PDU by resolving their values and produces a \texttt{Response} PDU. The \texttt{Response} PDU is sent back to the manager. If the processing of any variable binding fails the details of the error are reflected in the the error status and error index fields in the \texttt{Response} PDU. The \texttt{GetNext} operation retrieves the value of the next existing MIB instances in lexicographic order. Successive \texttt{GetNext} operations can be used to walk the MIB instances without prior knowledge about the MIB structure. The \texttt{GetBulk} operation is a generalization of the \texttt{GetNext} operation where the agent performs a series of \texttt{GetNext} operations internally. This makes it easier to acquire large amounts of related information without initiating repeated \texttt{GetNext} operations. The manager can set values for the managed objects by using the \texttt{Set} operation.

The \texttt{Trap} is sent by an agent to a manager and indicates that an event has occurred or a certain condition is present. There is no confirmation associated with this notification delivery mechanism. The confirmed notification delivery mechanism is provided by the \texttt{Inform} operation. Finally, the \texttt{Report} operation can be used internally to allow SNMP engines to communicate with each other for error notifications and clock synchronization.

\section{Security Models}
SNMPv1 and SNMPv2c  provide only trivial security based on a clear-text community string and lack of adequate mechanisms for securing the management function. The security problems are addressed in SNMPv3 where the User-based Security Model (USM) \cite{rfc3414} is defined that implements the security services for authentication and privacy. The USM provides three levels of security:
\begin{enumerate}
\item[1)] \texttt{noAuthNoPriv} mode provides no authentication and no encryption services.
\item[2)] \texttt{authNoPriv} mode provides message authentication, message integrity, and timeliness checking services but no encryption.
\item[3)] \texttt{authPriv} mode message provides authentication, message integrity, timeliness checking services and encryption of the payload.
\end{enumerate}

The document \cite{rfc5591} introduces a Transport Security Model for the SNMP protocol that makes use of existing and commonly deployed security infrastructures and Transport Models providing lower-layer, secure transports.
